Kronos \& Krios: Evidence for accretion of massive, rocky planetary system in a sun-like pair

### 2017-06-05 paper/ms.tex
Clean up discussion; comment out some non critical parts

% chance pair
Given that their metallicities and abundance patterns are significantly
different, one may simply conclude that the two stars are not related (coeval)
but they merely happen to be comoving at such small separation ($\approx 0.6$~pc)
by chance.
The two stars are in the Galactic disk, and assuming certain velocity ellipsoid
at the stars' location, one may compute the probability that a star moving at
the mean velocity of the two stars would have a companion within $\Delta v = XX$~km/s.
...

@@

% rocky planets (or the like) engulfment
Yet another more exotic possibility is that a star's surface abundance may
be altered by accretion of planetary material after birth.
Instabilities may develop in a multi-planet system due to its chaotic nature
which may lead to planet engulfment or ejection by the host star.
%TODO: needs references
% kozai?
Indeed, it is an important goal of many exoplanet studies
to detect chemical signatures of planet formation or accretion,
distinguish them from Galactic chemical evolution, and
connect them to theories of evolution of planetary systems.
One approach that is free from confusion with Galactic chemical evolution
is to compare two almost identical stars in a wide binary system,
which are assumed to be born together.

%One proxy that is informative of the composition of the material
%that was hypothetically accreted to a star is the condensation temperature (\Tcondens).
% TODO: why?

Only a handful of wide binary systems have been studied with high-resolution
spectroscopy in their detailed chemical abundances so far.
These systems are HD~20782/HD~20781, HD~80606/HD~80607, XO-2N/XO-2S, HAT-P-1,
WASP-94A/WASP94-B, and HD~133131A/HD~133131B.
We first review and summarize the studies of these systems focusing on key
characteristics of stars and planets, and interpretation of any trend in
$\Delta\elemH{X}$.
These studies commonly look for
%TODO: needs to include **16 Cyg** and HD~133131AB

@@

What about \elem{Li}?
The element \elem{Li} is worth a special attention in the context of accretion scenario.
Because Li is present in either carbonaceous chondrites or the bulk Earth with
a concentration of $1-1.5$~ppm in mass (\citealt{mcdonough2001composition}),
but is depleted quickly within the first couple of Gyrs on the surface of a
Sun-like star, accretion of either material at later times will significantly
replenish the lithium on the star's surface.
For the present-day Sun, the accretion of $20~\mearth$ of bulk Earth-like
material results in $\Delta\elemH{Li} \approx 1.6$~dex (which
is indicated as an upward arrow in \figname~\ref{fig:toycalc}).
Incidentally, the star under examination, \bizarreone, has an age (informed by
stellar parameters) very close to the Sun ($4.28_{-1.03}^{+1.11}$).
Thus, as long as we believe the Sun's present-day surface \elem{Li} abundance
to be ordinary, this predicts that the \bizarreone\ should have $1.6$~dex enhancement
in its \elem{Li}.
This is in fact exactly what we find: the \elem{Li} abundance of \bizarreone\ is
$A(\elem{Li}) = 2.75$ (Table~\ref{tab:kk}, \citealt{jmlithium})
approximately $1.65$~dex higher than the solar value of $1.1$~dex.
% TODO: solar Li needs reference
Furthermore, this implies that the accretion event should have happened very
recently from a time ago that is shorter than the \elem{Li} depletion time.

It is further worth pointing out that measuring the \elem{Li} abundance
with some of the refractory elements such as $\elem{Fe}$
provides a way to estimate the time of the event.
If the observed \elem{Li} abundance is less than what is inferred from
abundance differences of other elements, and
we have a model of \elem{Li} depletion, the time of \elem{Li} enhancement
may be calculated.

@@

We stress that while the calculation carried out is useful in
order-of-magnitude sense, further investigation of each of the simplifying
assumptions made is warranted.
In addition,

Given these limitations, the level of agreement for the refractory {\it and}
\elem{Li} for \bizarreone\ is remarkable.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{toycalc.png}
  \caption{Expected change in solar surface abundance after adding $20~\mearth$ of
    material with bulk Earth composition (\citealt{mcdonough2001composition}).
    The assumed mass fraction in the convective zone is $0.03$.
    The close match with the observed abundance difference in \bizarreone-\sunanalog\ pair
    suggests that the abundance difference may be due to accretion of
    $20~\mearth$ of rocky planetary material.
    We elements measured for \bizarreone-\sunanalog\ pair are filled with purple
    while those without measurements are left empty.
  }
  \label{fig:toycalc}
\end{figure}

@@

%Implications
%- chemical tagging, frequency is of concern
%- exoplaneteers looking for accretion signature: look for the most odd ones in
%the expected direction not just in binaries already known to host a planet.

>>>>

% chance pair
%Given that their metallicities and abundance patterns are significantly
%different, one may simply conclude that the two stars are not related (coeval)
%but they merely happen to be comoving at such small separation ($\approx 0.6$~pc)
%by chance.
%The two stars are in the Galactic disk, and assuming certain velocity ellipsoid
%at the stars' location, one may compute the probability that a star moving at
%the mean velocity of the two stars would have a companion within $\Delta v = XX$~km/s.
%...

@@

% rocky planets (or the like) engulfment
Yet another more exotic possibility is that a star's surface abundance may
be altered by accretion of planetary material after birth.
Instabilities may develop in a multi-planet system due to its chaotic nature
which may lead to planet engulfment or ejection by the host star.
%TODO: needs references
%TODO: kozai?
Indeed, it is an important goal of many exoplanet studies
to detect chemical signatures of planet formation or accretion,
distinguish them from Galactic chemical evolution, and
connect them to theories of evolution of planetary systems.
One approach that is free from confusion with Galactic chemical evolution
is to compare two almost identical stars in a wide binary system,
which are assumed to be born together.
Assuming that the component stars were born together with identical
initial composition, we may see a difference in their surface abundances
if the two stars then went through different amount of accretion
of planetary material.
This abundance difference may depend on the condensation temperatures of
elements in the protoplanetary disks from which the accreted planets formed
as their compositions depend on the radial temperature gradient in the disk.
%TODO: cautions on condensation temperature
% 1. it is just a proxy
% 2. it is **equlibrium** condensation temperature

Only a handful of wide binary systems have been studied with high-resolution
spectroscopy in their detailed chemical abundances so far.
These systems are HD~20782/HD~20781, HD~80606/HD~80607, XO-2N/XO-2S, HAT-P-1,
WASP-94A/WASP94-B, and HD~133131A/HD~133131B.
We first review and summarize the studies of these systems focusing on key
characteristics of stars and planets, and interpretation of any trend in
$\Delta\elemH{X}$ with \Tcondens.
%TODO: needs to include **16 Cyg** and HD~133131AB

@@

What about \elem{Li}?
The element \elem{Li} is worth a special attention in the context of accretion scenario.
Because Li is present in either carbonaceous chondrites or the bulk Earth with
a concentration of $1-1.5$~ppm in mass (\citealt{mcdonough2001composition}),
but is depleted quickly within the first couple of Gyrs on the surface of a
Sun-like star, accretion of either material at later times will significantly
replenish the lithium on the star's surface.
For the present-day Sun, the accretion of $20~\mearth$ of bulk Earth-like
material results in $\Delta\elemH{Li} \approx 1.6$~dex (which
is indicated as an upward arrow in \figname~\ref{fig:toycalc}).
Incidentally, the star under examination, \bizarreone, has an age (informed by
stellar parameters) very close to the Sun ($4.28_{-1.03}^{+1.11}$~Gyr).
Thus, as long as we believe the Sun's present-day surface \elem{Li} abundance
to be ordinary, this predicts that the \bizarreone\ should have $1.6$~dex enhancement
in its \elem{Li}.
This is in fact exactly what we find: the \elem{Li} abundance of \bizarreone\ is
$A(\elem{Li}) = 2.75$ (Table~\ref{tab:kk}, \citealt{jmlithium})
approximately $1.65$~dex higher than the solar value of $1.1$~dex.
%TODO: solar Li needs reference
Furthermore, this implies that the accretion event should have happened very
recently from a time ago that is shorter than the \elem{Li} depletion time.

%TODO: not sure if this is too obvious of a point or useful in practice..
%It is further worth pointing out that measuring the \elem{Li} abundance
%with some of the refractory elements such as $\elem{Fe}$
%provides a way to estimate the time of the event.
%If the observed \elem{Li} abundance is less than what is inferred from
%abundance differences of other elements, and
%we have a model of \elem{Li} depletion, the time of \elem{Li} enhancement
%may be calculated.

@@

%TODO: augment this paragraph
% - more geochemical study references
We stress that while the calculation carried out is useful in
order-of-magnitude sense, further investigation of each of the simplifying
assumptions made is warranted.
In addition, the composition of bulk Earth also has some uncertainties.
For example, the reported bulk Earth concentration of the siderphile element
\elem{Mn}, varies from $800$ to $1700$~ppm (\citealt{1998psc..book.....L}).
While the latter value from \citealt{mcdonough2001composition} has been used in
our calculation, the former value would bring the observed $\Delta\elemH{Mn}$
to an even closer agreement.
Given these limitations, the level of agreement for the refractory {\it and}
\elem{Li} for \bizarreone\ is remarkable.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{toycalc.png}
  \caption{
    Comparing the observed abundance difference vs. \Tcondens\ rank
    to the expected change in solar surface abundance after adding $20$~\mearth\ of
    material with bulk Earth composition (\citealt{mcdonough2001composition}).
    The assumed mass fraction in the convective zone is $0.03$.
    All metals are ranked by their \Tcondens\ for solar composition gas,
    same as in \figname~\ref{fig:relabun_tcrank}.
    Locations of the elements measured for \bizarreone-\sunanalog\ pair are
    indicated by a vertical line and its symbol.
    The close match with the observed abundance difference in \bizarreone-\sunanalog\ pair
    suggests that the abundance difference may be due to accretion of
    $20$~\mearth\ of rocky planetary material.
    The element \elem{Li} is off the plot and indicated with a red arrow
    (see text for details).
  }
  \label{fig:toycalc}
\end{figure}

@@

%TODO: Implications
%- chemical tagging, frequency is of concern
%- exoplaneteers looking for accretion signature: look for the most odd ones in
%the expected direction not just in binaries already known to host a planet.
